Job mobility and earnings growth: diffrerent markets, different outcomes.

### 2018-09-14 draft.Rmd
Update draft.Rmd

Countries differ in their rate of job mobility and turnover. Researchers cite employment protection legislation as key to understanding the rates; however, the rewards or effects tied to mobility play a stronger role in determining differences. We consider this idea through two aims. First we compare the effects of quits and promotions on earnings growth. Second, we consider their effects in two institutional settings (Germany and the UK). Studies exploring the effects of job mobility are rare, studies which compare and contrast their effects in separate institutional frameworks, or capitalisms, are rarer. We offer two findings. First, quits and promotions have significantly different effects on earnings growth, depsite both commonly labeled as "voluntary". Internal promotions yield significant earnings growth for workers in both countries. Quits have a minor effect, which is explained by a change in weekly working hours. Second, the effect of promotions is larger in Germany than in the UK. This suggests the premium tied to promotion is greater in Germany than the UK. The high premium tied to internal mobility in Germany may act as an incentive to commit to the firm instead of the market, hence the lower rate of mobility among German workers. The article helps to understand country-differences in mobility rates; our explanations go beyond a nation's level of employment protection.

>>>>

**Purpose:** This article compares internal and external job mobility as a means of improving earnings growth in Germany and the UK.

**Design/methodology/approach:** We present parallel case studies using the British Household Panel Survey, and the German Socio-Economic Panel. Within both panels, we consider job history questions which operationalise voluntary internal and external job mobility. Using this data we estimate the effect of job mobility on earnings growth using fixed-effects estimation, and compare both countries.

**Findings:** Internal and external mobility are distinctly different strategies for improving earnings. However, the institutional setting under which mobility takes place also shapes the rewards tied to mobility. In both countries the core workforce benefits form internal promotions. However, in Germany the rewards tied to internal promotion are far greater than they are in the UK. One of the reasons German workers are less mobile than British workers may have less to do with employment protection legislation and more to do with the high rewards tied to internal career ladders.

**Practical implications:**

Countries differ in their rates of job mobility and turnover. Researchers cite employment protection legislation as key to understanding the rates; however, the rewards or effects tied to mobility play a stronger role in determining differences. We consider this idea through two aims. First we compare the effects of quits and promotions on earnings growth. Second, we consider their effects in two institutional settings (Germany and the UK). Studies exploring the effects of job mobility are rare, studies which compare and contrast their effects in separate institutional frameworks, or capitalisms, are rarer. We offer two findings. First, quits and promotions have significantly different effects on earnings growth, depsite both commonly labeled as "voluntary". Internal promotions yield significant earnings growth for workers in both countries. Quits have a minor effect, which is explained by a change in weekly working hours. Second, the effect of promotions is larger in Germany than in the UK. This suggests the premium tied to promotion is greater in Germany than the UK. The high premium tied to internal mobility in Germany may act as an incentive to commit to the firm instead of the market, hence the lower rate of mobility among German workers. The article helps to understand country-differences in mobility rates; our explanations go beyond a nation's level of employment protection.
